Ceph Parallel File System Evaluation Report

### 2013-08-16 ceph-eval.tex
apply scott's patch

\usepackage{fontspec}
\usepackage{mathptm}
\usepackage{comment}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{fancyvrb}
\usepackage{float}
\usepackage{setspace}

@@

\title{Ceph Parallel File System Evaluation Report}
\author{Feiyi Wang, Sarp Oral, Doug Fuller, Blake Caldwell, James Simmons\\
Brad Settlemyer, Scotty Atchley, Jason Hill, Mark Nelson}

@@

NCCS contracted Inktank, the major developer behind the open-sourced parallel
file system, Ceph, for a deep-dive study. It covers a wide range of technical
topics: from system setup, deployment, administration to internal mechanism. In
comparison to other parallel file systems, Ceph has a number of distinctive
features:

\begin{itemize}
  \item Ceph has an intelligent and powerful data placement mechanism, known as
  CRUSH. CRUSH algorithm allows the client side to pre-calculate the layout
  info, and take failure domain, hierarchical storage tiers into placement
  consideration.

  \item From the on start, Ceph is designed to manage metadata and name space
  with a cluster of metadata servers.

\item Ceph is designed with the assumption that system is composed with
unreliable components, so replication and failure detection and tolerance are
the norm rather than the exception. This is in line with the expectation of
exascale computing.

@@

\item Most of Ceph code base are in user-space. Generally speaking, it makes the
system easier to debug and maintain. The client side support has long been
integrated into Linux mainline kernel, which make the deployment and out-of-box
experience easier.

@@

As part of this effort, we set up a dedicated testbed within NCCS for Ceph
system evaluation. The goal is to investigate the feasibility of Ceph system as
a potential candidate for our next-generation HPC storage solution.
This report presents our experience, results, and observations. While evaluating
our results, please keep in mind that Ceph is such a moving target that its
code base is changing rapidly. In between releases, we often see different
outcomes. We will try to make clear when such change happens in the report.

@@

The storage backend we have for this evaluation is DDN's SFA10K. It consists of
200 SAS drives and 280 SATA drives, driven by 4 hosts. Each host has 2 IB QDR
connections to the storage backend. We configure and use 4 Mellanox IB cards as
primary connections, which by calculation can saturate the maximum theoretical
throughput, around 11 - 12 GB/s. The connection diagram is illustrated as the
following:

@@

The hosts are running Redhat 6.3 with 3.5.1 kernel (rhl-ceph image), Glibc 2.12
with syncfs support, locally patched. The Ceph version is 0.55 release in the
initial test, then upgraded to 0.64 and 0.67RC for second round of test. For a
complete a list of hosts that are running ceph images, you can do:

\begin{Verbatim}
grep "rhel6-ceph" /etc/gedi/MAC.info
\end{Verbatim}

@@

\subsection*{Block I/O}
We first establish the baseline performance by measuring block I/O performance.
At the block level, with LUN configured as RAID6 8+2, we have the following
results:

\begin{table}[htb]
\centering
\begin{tabular}{ll}
    \toprule
    SAS single LUN sequential read & ~1.4 GB/s \\
    SATA single LUN sequential read & ~955 MB/s \\[0.5em]
    Single host with 4 SAS LUNs & ~ 2.8 GB/s \\
    Single host with 7 SATA LUNs & ~ 2.6 GB/s \\
    \bottomrule
\end{tabular}

@@

Single host refers to one of 4 tick-oss nodes, which drives the entire SSU.
Overall, the entire SSU performance can perform at 11GB/s, comparing to
theoretical maximum of 12GB/s.

@@

Note that we are using IP over IB for networking, since Ceph doesn't have
native IB support\footnote{Inktank is developing rsocket alternative to better take
advantage of IB fabric}.
A single host with IP over IB (QDR) can perform at 2.7 GB per second with
Mellanox recommended tuning applied. With all 4 hosts (OSD servers), we expect
the aggregate throughput is in the neighborhood of 10GB/s.

Unfortunately there was not enough time to do more detailed analysis of network
bisection bandwidth. Later, we will see RADOS is performing a
maximum of 5 to 6 GB per seconds driven by 4 hosts.  So, this test shows that we
have enough network bandwidth, i.e., IP over IB is not a bottleneck in this
case.

@@

As we plan to use XFS file system as the backend local file system for Ceph, we
examine XFS against SFA10K to acquire a set of sensible configuration parameters
for optimized performance. We sampled a selected set of parameters (block size,
queue depth, request size, sequential read and write), and this is the
configurations we selected: mount with \verb!nobarrier,noatim,inode64! options.
The \verb!inode64! option had a notable improvement on sequential write (around
20\%).

\begin{figure}[htb]
\centering
%% -- 1st figure
\begin{minipage}[t]{0.5\linewidth}
\centering
\includegraphics[width=3in]{data/xdd-read}
\caption{XFS read scaling on \# of devices}
\label{fig:xfs-read}
\end{minipage}%
\begin{minipage}[t]{0.5\linewidth}
\centering
\includegraphics[width=3in]{data/xdd-write}
\caption{XFS write scaling on \# of devices}
\label{fig:xfs-write}
\end{minipage}%
\end{figure}

The read and write result are summarized in Figure~\ref{fig:xfs-read}
and Figure~\ref{fig:xfs-write}. These scaling test on the number of devices
indicates that XFS can get to peak read performance with 5 SATA LUNs, then there
is a bit of degradation from there. The parameter space explored regarding to
the queue depth doesn't seem to make any obvious difference on the performance
though.

@@

RADOS is Ceph object store, a foundational component for CephFS file
system. There are three types of scaling tests we are interested at RADOS layer:

@@

Our system setup poses some limitation on the scalability test we want to
perform. In particular, we currently have 4 OSS server, 8 clients, and 11 OSD
per clients. So the scaling test will be within these constraints.

An open source tool developed by Inktank called RADOS bench was do perform
initial performance analysis of the underlying RADOS layer.  RADOS Bench simply
writes out objects to the underlying object storage as fast as possible, and
then later reads those objects in the same order as they were written.

We've observed that using more than one client processes and lots of concurrent
operations is important when performing these tests.  8 Client processes are
used with 32 concurrent 4MB Objects in flight each, and a pool is created for
each RADOS bench process to ensure that object reads come from independent pools
(RADOS bench is not smart enough to ensure that objects are not read by multiple
processes and thus possibly cached).  A sync and flush is performed on every
node before every test to ensure no data or metadata is in cache.  All tests
were run with 1X replication.  Unfortunately at this time only XFS was present
and usable for the underlying OSDs so BTRFS and EXT4 file systems were not
tested at this time.

@@

In the following test, a single Ceph host drives $n$ number of OSDs, where $n$
increases from 1 to 11. The result is illustrated in Figure~\ref{fig:osd-scale}.
We ran the test against a single client with 4 concurrent process. In this test
case, we observe that Ceph OSS exhibits near linear scalability up to 9
OSDs, and still have room to grow at 11 OSDs. This suggests that we have not reached the
saturation point yet. It requires provisioning of more OSDs at SFA10K backend.

@@

In this test, we exercise OSS number from 1 to 4, driven by 4 clients with 4
concurrent process for each client. Each addition of OSS adds 11 more OSDs into
the play. We observe that Ceph exhibits linear scaling property with regard to
number of servers as well, at least in the given set of servers. However, the peak
performance we are seeing is about the half of what are expecting from SFA10K.
For the write, the reason behind it is attributed to the way Ceph perform
journaling: Ceph doesn't support the meta-data only journaling, therefore every
write is equivalent of a double-write: once to the data device, once to the
journaling device. It effectively cut the system bandwidth in half. That said,
it doesn't explain the read performance - though it is better than write, but is
still far from the theoretical maximum.

@@

\begin{table}[tb]
\centering
\begin{tabular}{p{1.5in} | p{3in}}
    \toprule
    IOR parameter & Note \\ \midrule
    \verb!-F! & file per process \\ \midrule
    \verb!-a POSIX! & use POSIX API \\ \midrule
    \verb!-w -r -C! & do both write and read test, \verb!-C! is to change task
        ordering for read back so it won't read from the write cache. \\ \midrule
    \verb!-i 3 -d 5! & 3 iterations and delay 5 seconds betewen iterations \\
    \midrule  
    \verb!-e! & perform \verb!fsync()! upon POSIX write close \\ \midrule
    \verb!-b 8g or 16g! & the block size \\ \midrule
    \verb!-t 4k to 4m! & the transfer size \\ \midrule
    \verb!-o file! & mandatory test file  \\    
    \bottomrule
\end{tabular}
\caption{IOR parameter setup}
\label{tbl:ior}
\end{table}

@@

The full permutation of IOR parameters are not explored due to IO errors we ran
into. We were, however, able to record results in two extreme case as far as
transfer size is concerned: 4KB and 4MB, using the same number of OSS servers
(4) and same block size (8GB), the results are illustrated in Figure
\ref{fig:ior4k} and \ref{fig:ior4m}, we make the following observations:

@@

With cache mirroring disabled write performance when using all 4 servers
improved dramatically, as illustrated in
Figure~\ref{fig:rados-ddn-mirror-disabled}. With BTRFS for example, we hit over
5.5GB/s from the clients.  When accounting for journal writes, that is over
11GB/s to the disks and very close to what the DDN chassis is capable of doing. 
Unfortunately, read performance did not scale as well.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=5in]{rados-064-osd}
\caption{RADOS Bench Scaling on \# of OSD, Ceph 0.64, 4MB IO, 8 Client Nodes}
\label{fig:rados-064-osd}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=5in]{rados-064-oss}
\caption{RADOS Bench Scaling on \# of Server, Ceph 0.64, 4MB IO, 8 Client Nodes}
\label{fig:rados-064-oss}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{ior-client-no-crc32}
\caption{IOR test with no client side CRC32}
\label{fig:ior-no-client-crc32}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{ior-kernel-39}
\caption{Kernel changes to 3.9, IOR with 4MB transfer size}
\label{fig:ior-kernel-39}
\end{figure}

@@

As before, we ran IOR scaling test with two cases: transfer size 4K and 4M.
These results are illustrated in Figure~\ref{fig:ior-064}. As expected, we are
seeing a much improved picture of read and write performance, very much in
line with RADOS bench performance.

\begin{figure}[htb]
\centering
\includegraphics[width=5in]{ior-064-4k}
\includegraphics[width=5in]{ior-064-4m}
\caption{IOR Scaling Test: 4KB and 4MB transfer size}
\label{fig:ior-064}
\end{figure}

We observe that a couple of places where the average write throughput is quite a
bit lower than the max.  This behavior was observed during other tests as well,
indicating that we may have periods of time where write throughput is
temporarily degrading.  Despite these issues, performance general seems to be
improving as the number of clients increases.  Writes seem to be topping out at
around 5.2GB/s (which is roughly what we would expect).  Reads seem to be
topping out anywhere from 5.6-7GB/s, however it is unclear if read performance
would continue scaling with more clients and get closer to the 8GB/s+ numbers we
saw with RADOS Bench.

@@

\item \verb!-n 1000!: this is per client file \textit{and} directories. For 8
clients, the number of files and directories is 8,000. Since we didn't specify
either \verb!-D! or \verb!-F!, so this is a mixture of both.

@@

Each test iterates 5 times and we are showing the max. 
We summarize the results as follows:

@@

  \item Ceph is built on the assumption that the underlying hardware components
  is unreliable, with little or no redundancy and failure detection capability.
  This is not the case in this leadership computing facilities. We have disabled
  replication for pools, but this will take a toll somewhere in the processing
  and we don't know if this is significant.

>>>>

\usepackage{fontspec}
\usepackage{mathptm}
\usepackage{comment}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{fancyvrb}
\usepackage{float}
\usepackage{setspace}
\usepackage{color,soul}

@@

\title{Ceph Parallel File System Evaluation Report}
\author{Feiyi Wang, Sarp Oral, Doug Fuller, Blake Caldwell, James Simmons\\
Brad Settlemyer, Scott Atchley, Jason Hill, Mark Nelson}

@@

NCCS contracted Inktank, the major developer behind the open-sourced parallel
file system, Ceph, for a deep-dive study. It covers a wide range of technical
topics: from system setup, deployment, administration to \hl{internal mechanism}. In
comparison to other parallel file systems, Ceph has a number of distinctive
features:

\begin{itemize}
  \item Ceph has an intelligent and powerful data placement mechanism, known as
  CRUSH. The CRUSH algorithm allows a client to pre-calculate the layout
  info, while taking into consideration failure domains and hierarchical storage tiers.

  \item From the start, Ceph's design anticipated managing metadata and the name space
  with a cluster of metadata servers.

\item Ceph's design assumes that the system is composed of
unreliable components; fault-detection and fault-tolerance (e.g., replication) are
the norm rather than the exception. This is in line with the expectation of
exascale computing.

@@

\item Most of the Ceph processes reside in user-space. Generally speaking, it makes the
system easier to debug and maintain. The client-side support has long been
integrated into Linux mainline kernel, which make the deployment and out-of-box
experience easier.

@@

As part of this effort, we set up a dedicated testbed within NCCS for Ceph
system evaluation. The goal is to investigate the feasibility of using the Ceph system as
a potential candidate for our next-generation HPC storage solution.
This report presents our experience, results, and observations. While evaluating
our results, please keep in mind that Ceph is such a moving target that its
code base is changing rapidly. In between releases, we often experienced different
outcomes. We will try to make clear when such changes happened in the report.

@@

We used Data Direct Network's (DDN) SFA10K as the storage backend during this evaluation. It consists of
200 SAS drives and 280 SATA drives, driven by four hosts. Each host has two InfiniBand (IB) QDR
connections to the storage backend. We configured and used four Mellanox IB cards as the
primary connections, which by calculation can saturate the maximum theoretical
throughput (11-12 GB/s). The connection diagram is illustrated as the
following:

@@

The hosts ran Redhat 6.3 with 3.5.1 kernel (rhl-ceph image), Glibc 2.12
with syncfs support, locally patched. We used the Ceph 0.55 release in the
initial tests, then upgraded to 0.64 and then to 0.67RC for second round of tests. For a
complete a list of hosts that are running ceph images, one can execute:

\begin{Verbatim}
$ grep "rhel6-ceph" /etc/gedi/MAC.info
\end{Verbatim}

@@

\subsection*{Block I/O}
We first establish the baseline performance by measuring block I/O performance.
At the block level, with each LUN configured with RAID6 8+2, we have the following
results:

\begin{table}[htb]
\centering
\begin{tabular}{ll}
    \toprule
    SAS single LUN sequential read & ~1.4 GB/s \\
    SATA single LUN sequential read & ~955 MB/s \\[0.5em]
    Single host with four SAS LUNs & ~ 2.8 GB/s \\
    Single host with seven SATA LUNs & ~ 2.6 GB/s \\
    \bottomrule
\end{tabular}

@@

Single host refers to one of four tick-oss nodes, which drives the entire SSU.
Overall, the entire SSU performance can perform at 11 GB/s, comparing to
theoretical maximum of 12 GB/s.

@@

Note, Ceph uses the BSD Sockets interface in the \texttt{SOCK\_STREAM} mode (i.e., TCP).
Because our testbed has IB, we used IP over IB (IPoIB) for networking\footnote{Inktank is 
investigating using rsockets to improve performance when using an IB fabric}.
A pair of hosts connected by IB QDR using IPoIB can transfer data at 2.7 GB/s if the
hosts are tuned per recommendations from Mellanox. With all four hosts (OSD servers), we would expect
the aggregate throughput to be in the neighborhood of 10 GB/s.

Unfortunately there was not enough time to do more detailed analysis of the network
performance. Later, we will see RADOS is performing a
maximum of 5-6 GB/s driven by four hosts.  So, this test shows that we
have enough network bandwidth (i.e., IP over IB is not a bottleneck in this
case).

@@

Ceph supports multiple backend file systems including btrfs, ext4, and xfs.
Due to its performance and maturity, we chose to use XFS. We
experimented with XFS to acquire a set of sensible configuration parameters
which hopefully provided optimal performance for the SFA10K. We sampled a selected set of parameters (block size,
queue depth, request size, sequential read and write). We settled on the following
configuration: mount with \verb!nobarrier,noatim,inode64! options.
The \verb!inode64! option had a notable improvement on sequential write (around
20\%).

%% Can the graphs be adjusted to have the same values in the Y axis?
\begin{figure}[htb]
\centering
%% -- 1st figure
\begin{minipage}[t]{0.5\linewidth}
\centering
\includegraphics[width=3in]{data/xdd-read}
\caption{XFS read scaling on \# of devices}
\label{fig:xfs-read}
\end{minipage}%
\begin{minipage}[t]{0.5\linewidth}
\centering
\includegraphics[width=3in]{data/xdd-write}
\caption{XFS write scaling on \# of devices}
\label{fig:xfs-write}
\end{minipage}%
\end{figure}

%% Are the read and write graphs wrong or should the second sentence below be about write
%% instead of read?
The read and write results are summarized in Figures~\ref{fig:xfs-read}
and~\ref{fig:xfs-write}, respectively. These show scaling behavior over the number of devices, which
indicates that XFS can get to the peak \hl{read} performance with five SATA LUNs, then there
is a bit of degradation from there. Changes to
the queue depth do not seem to make any obvious difference on the performance.

@@

RADOS is the Ceph object store, a foundational component for CephFS file
system. There are three types of scaling tests we are interested at RADOS layer:

@@

Our system setup poses some limitation on the scalability tests we want to
perform. In particular, we currently have four OSS servers, eight clients, and eleven OSD servers
per client. The scaling tests will be within these constraints.

We use the open-source tool RADOS bench, developed by Inktank, to perform our
initial performance analysis of the underlying RADOS layer.  RADOS Bench simply
writes out objects to the underlying object storage as fast as possible, and
then later reads those objects in the same order as they were written.

We have observed that using two or more client processes and many concurrent
operations are important when performing these tests.  We tested eight client processes 
with 32 concurrent 4 MB objects in flight each. We created a pool for
each RADOS bench process to ensure that object reads come from independent pools
(RADOS bench is not smart enough to ensure that objects are not read by multiple
processes and thus possibly cached).  A sync and flush is performed on every
node before every test to ensure no data or metadata is in cache.  All tests
were run with replication set to one.  Only XFS was present
and usable for the underlying OSDs, so BTRFS and EXT4 file systems were not
tested at this time.

@@

In the following test, a single Ceph host drives $n$ number of OSDs, where $n$
increases from one to eleven. The result is illustrated in Figure~\ref{fig:osd-scale}.
We ran the test against a single client with four concurrent processes. In this test
case, we observe that the OSS exhibits near linear scalability up to nine
OSDs, and is still trending upwards at eleven OSDs. This suggests that we have not reached the
saturation point yet. Additional testing would require provisioning more OSDs on the SFA10K backend.

@@

In this test, we exercise OSS number from one to four, driven by four clients with four
concurrent process for each client. Each addition of OSS adds eleven more OSDs into
play. We observe that Ceph exhibits linear scaling property with regard to
number of servers as well, at least in the given set of servers. However, the peak
performance we are seeing is about the half of what are expecting from SFA10K.
For the write, the reason behind it is attributed to the way Ceph perform
journaling: Ceph does not support the meta-data only journaling, therefore every
write is equivalent of a double-write: once to the data device, once to the
journaling device. It effectively cut the system bandwidth in half. That said,
it does not explain the read performance - though it is better than write, but is
still far from the theoretical maximum.

@@

\begin{table}[tb]
\centering
\begin{tabular}{p{1.5in} | p{3in}}
    \toprule
    IOR parameter & Note \\ \midrule
    \verb!-F! & file per process \\ \midrule
    \verb!-a POSIX! & use POSIX API \\ \midrule
    \verb!-w -r -C! & do both write and read test, \verb!-C! is to change task
        ordering for read back so it will not read from the write cache. \\ \midrule
    \verb!-i 3 -d 5! & 3 iterations and delay 5 seconds betewen iterations \\
    \midrule  
    \verb!-e! & perform \verb!fsync()! upon POSIX write close \\ \midrule
    \verb!-b 8g or 16g! & the block size \\ \midrule
    \verb!-t 4k to 4m! & the transfer size \\ \midrule
    \verb!-o file! & mandatory test file  \\    
    \bottomrule
\end{tabular}
\caption{IOR parameter setup}
\label{tbl:ior}
\end{table}

@@

The full permutation of IOR parameters are not explored due to IO errors we ran
into. We were, however, able to record results in two extreme case as far as
transfer size is concerned: 4 KB and 4 MB, using the same number of OSS servers
(4) and same block size (8 GB), the results are illustrated in Figure
\ref{fig:ior4k} and \ref{fig:ior4m}, we make the following observations:

@@

With cache mirroring disabled write performance when using all four servers
improved dramatically, as illustrated in
Figure~\ref{fig:rados-ddn-mirror-disabled}. With BTRFS for example, we hit over
5.5 GB/s from the clients.  When accounting for journal writes, that is over
11 GB/s to the disks and very close to what the DDN chassis is capable of doing. 
Unfortunately, read performance did not scale as well.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=5in]{rados-064-osd}
\caption{RADOS Bench Scaling on \# of OSD, Ceph 0.64, 4 MB IO, 8 Client Nodes}
\label{fig:rados-064-osd}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=5in]{rados-064-oss}
\caption{RADOS Bench Scaling on \# of Server, Ceph 0.64, 4 MB IO, 8 Client Nodes}
\label{fig:rados-064-oss}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{ior-client-no-crc32}
\caption{IOR test with no client-side CRC32}
\label{fig:ior-no-client-crc32}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{ior-kernel-39}
\caption{Kernel changes to 3.9, IOR with 4 MB transfer size}
\label{fig:ior-kernel-39}
\end{figure}

@@

As before, we ran IOR scaling test with two cases: transfer size 4 KB and 4 MB.
These results are illustrated in Figure~\ref{fig:ior-064}. As expected, we are
seeing a much improved picture of read and write performance, very much in
line with RADOS bench performance.

\begin{figure}[htb]
\centering
\includegraphics[width=5in]{ior-064-4k}
\includegraphics[width=5in]{ior-064-4m}
\caption{IOR Scaling Test: 4 KB and 4 MB transfer size}
\label{fig:ior-064}
\end{figure}

We observe that a couple of places where the average write throughput is quite a
bit lower than the max.  This behavior was observed during other tests as well,
indicating that we may have periods of time where write throughput is
temporarily degrading.  Despite these issues, performance general seems to be
improving as the number of clients increases.  Writes seem to be topping out at
around 5.2 GB/s (which is roughly what we would expect).  Reads seem to be
topping out anywhere from 5.6-7 GB/s, however it is unclear if read performance
would continue scaling with more clients and get closer to the 8 GB/s+ numbers we
saw with RADOS Bench.

@@

\item \verb!-n 1000!: this is per client file \textit{and} directories. For eight
clients, the number of files and directories is 8,000. Since we did not specify
either \verb!-D! or \verb!-F!, so this is a mixture of both.

@@

Each test iterates five times and we are showing the max. 
We summarize the results as follows:

@@

  \item Ceph is built on the assumption that the underlying hardware components
  is unreliable, with little or no redundancy and failure detection capability.
  This is not the case in this leadership computing facilities. We have disabled
  replication for pools, but this will take a toll somewhere in the processing
  and we do not know if this is significant.
